---
title: "Disease_prediction"
author: "María Bueno Álvez"
date: "08/11/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

Analysis date: `r Sys.Date()`

# Set up

## Load libraries

```{r}

library(tidyverse)
library(ggbeeswarm)
library(ggrepel)
library(caret)
library(pROC)
library(patchwork)
library(gridExtra)
library(multiROC)
library(pheatmap)
library(ggplotify)
library(tidygraph)
library(ggraph)
library(impute)
library(limma)
library(writexl)

select <- dplyr::select
summarise <- dplyr::summarise
rename <- dplyr::rename

source("scripts/functions_utility.R")
source("scripts/functions_classifiers.R")

```

## Load data

```{r}

# Cancer data (synthetic)
data <-
  readRDS("data/cancer_data_synthetic.rds")
metadata <-
  readRDS("data/cancer_metadata_synthetic.rds")

# Healthy data (synthetic)
wellness_data <-
  readRDS("data/wellness_data_synthetic.rds")
wellness_meta <-
  readRDS("data/wellness_metadata_synthetic.rds")

# Themes & palettes
palettes <- 
  readRDS("data/processed/others/palettes.rds")
themes <- 
  readRDS("data/processed/others/plot_themes.rds")
levels <- 
  readRDS("data/processed/others/levels.rds")

```

# Split train - test

```{r}

cancer_data <-
  data %>% 
  select(Sample) %>% 
  distinct() %>% 
  left_join(metadata, by = "Sample") %>% 
  filter(GROUP %in% levels$cancers_12) %>% 
  select(Sample,GROUP)

cancers_split <-
  cancer_data %>% 
  group_by(GROUP) %>% 
  do({
    
    dat <- .
    
    set.seed(213)
    train <- createDataPartition(dat$GROUP,p=0.7,list=FALSE)
    train_data <- 
      dat[train,] %>% 
      mutate(set = "train")
    test_data <- dat[-train,] %>% 
      mutate(set = "test")
    train_data %>% 
      bind_rows(test_data)
  })
 
saveRDS(cancers_split, savepath_data("Data_split", "cancers_split.rds"))

```


# Disease classifier for the 12 cancers

## glmnet algorithm

```{r}

# Iterate over the 12 cancers to 1) generate a balanced set of controls (with similar number of patients as the cancer group) and 2) Build a disease classifier 
glmnet_res <- 
  lapply(levels$cancers_12, function(cancer) {
    
    cancer_data_split <- 
      generate_balanced_controls(cancer = cancer,
                                 metadata = metadata, 
                                 set_data = cancers_split,
                                 n_control_groups = length(levels$cancers_12)-1)
  
    disease_classifier(cancer_type = cancer,
                       method = "glmnet", 
                       data = data,
                       split_data = cancer_data_split,
                       metadata = metadata, 
                       select_proteins = NULL, 
                       tune_param = 10, 
                       CV = 5)
  }) %>% 
  set_names(levels$cancers_12)

saveRDS(glmnet_res, savepath_data("Models", "glmnet_res.rds"))

```


### Heatmap - proteins with > 25% importance in one cancer

```{r}

# Extract importance scores for each protein across all cancers
all_prot <- 
  lapply(c(1:length(levels$cancers_12)), function(i){
    model <- glmnet_res[[i]]
    name <- levels$cancers_12[[i]]
    
    model$prot_res %>% 
      mutate(GROUP = name)
  })

# Number of proteins with >25% importance per cancer
n_prots <- 
  all_prot %>% 
  bind_rows() %>% 
  rename(Cancer = GROUP) %>% 
  filter(Overall > 25) %>% 
  group_by(Cancer) %>% 
  summarise(n = n_distinct(Protein)) %>% 
  arrange(-n) 

n_prots %>% 
  mutate(Cancer = factor(Cancer, levels = n_prots$Cancer)) %>% 
  ggplot(aes(Cancer,n,fill=Cancer)) +
  geom_bar(stat="identity", show.legend = F) +
  geom_text(aes(label=n, y = n+15, color = Cancer), size = 3.5,  show.legend = F, color = "black") +
  themes$simple +
  scale_fill_manual(values = palettes$alt_group[levels$cancers_12]) +
  scale_color_manual(values = palettes$alt_group[levels$cancers_12]) 

ggsave(savepath("barplot_num_proteins_25_importance.pdf"),
       height = 2,
       width = 4)

# Importance across all cancers for proteins with >25% importance in at least one cancer 
prots <- 
  all_prot %>% 
  bind_rows() %>% 
  rename(Cancer = GROUP) %>% 
  filter(Overall > 25) %>% 
  pull(Protein) %>% 
  unique()

order <-
  tibble(GROUP = n_prots$Cancer) %>% 
  left_join(levels$cancers_12_mapping, by = "GROUP") %>% 
  pull(Cancer)

ann <- 
  palettes$alt_group_complete %>% 
  enframe() %>% 
  mutate(value = name) %>% 
  rename(Cancer = value) %>% 
  column_to_rownames("name")

all_prot %>% 
  bind_rows() %>% 
  left_join(levels$cancers_12_mapping, by = "GROUP") %>% 
  mutate(Cancer = factor(Cancer, levels = order)) %>% 
  filter(Protein %in% prots) %>% 
  select(-GROUP) %>% 
  spread(Cancer,Overall) %>% 
  column_to_rownames("Protein") %>% 
  pheatmap(color = colorRampPalette(c("grey95", "grey15"))(100),
           show_rownames = F, 
           cluster_cols = F,
           annotation_col = ann,
           annotation_colors = list("Cancer" = palettes$alt_group_complete)) %>% 
  as.ggplot() 

ggsave(savepath("glmnet_protein_importance_25.pdf"),
       height = 6,
       width = 4)

```

### Network - proteins with > 25% importance in one cancer

```{r}

# Prepare input data
network_model_data <-
  all_prot %>% 
  bind_rows() %>% 
  rename(Cancer = GROUP) %>% 
  filter(Overall > 25) %>% 
  select(Protein,Cancer,Overall)

node_pal <-
  c(unique(network_model_data$Protein),unique(network_model_data$Cancer)) %>%
  enframe() %>%
  mutate(name = value,
         value = case_when(name %in% levels$cancers_12 ~ palettes$alt_group[name],
                           T ~ "grey70")) %>%
  deframe()

# Generate network
set.seed(213)
network_model_data %>% 
  as_tbl_graph() %>% 
  activate(nodes) %>% 
  mutate(s = case_when(name %in% levels$cancers_12 ~ 5,
                       T ~ 0.5),
         text_color = case_when(name %in% levels$cancers_12 ~ "white",
                                T ~ "black")) %>%
  create_layout(layout = "nicely") %>% 
  ggraph() + 
  geom_edge_link(aes(), color = "grey90") + # Indicate width  = Overall for link proportional to importance score
 # scale_edge_width(range = c(0.1, 2)) +
  geom_node_point(aes(color = name, 
                      size = s),
                  show.legend = F)+
  geom_node_point(data = . %>% 
                    filter(name %in% levels$cancers_12) ,
                  aes(color = name,
                      size = s),
                  fill ="grey30",
                  shape = 21,
                  stroke = 3,
                  show.legend = F) +
  scale_color_manual(values = node_pal) +
  geom_node_text(data = . %>% 
                   filter(name %in% levels$cancers_12) ,
                 aes(label = str_wrap(name, width = 10)),
                 color = "white",
                 lineheight = 0.8,
                 size = 1.5,
                 vjust = 0,
                 hjust = 0.5) +
  theme_void() + 
  coord_fixed() +
  ggtitle("Overall > 25")
  
ggsave(savepath("network_proteins_25_importance.pdf"))

```

## Random forest (RF) algorithm

```{r}

# Iterate over the 12 cancers to 1) generate a balanced set of controls (with similar number of patients as the cancer group) and 2) Build a disease classifier 
rf_res <- 
  lapply(levels$cancers_12, function(cancer) {
    
    cancer_data_split <- 
      generate_balanced_controls(cancer = cancer,
                                 metadata = metadata,
                                 set_data = cancers_split,
                                 n_control_groups = length(levels$cancers_12)-1)
    
    disease_classifier(cancer_type = cancer,
                       method = "rf", 
                       data = data,
                       split_data = cancer_data_split,
                       metadata = metadata, 
                       select_proteins = NULL, 
                       tune_param = 10, 
                       CV = 5)
  }) %>% 
  set_names(levels$cancers_12)

saveRDS(rf_res, savepath_data("Models", "rf_res.rds"))

```

# Compare glmnet & RF results

## De-long test

```{r}

# Calculate whether the differences in AUC for the glmnet and RF models is significant using the De-long test 
delong_rf_glmnet <- 
  data.frame(cancer = levels$cancers_12) %>% 
  group_by(cancer) %>% 
  do({
    
    c <- .$cancer
    res <- 
      pROC::roc.test(glmnet_res[[c]]$roc_res,
                     rf_res[[c]]$roc_res, 
                     method = "delong", 
                     boot.n=10000)
    
    data.frame(AUC_glmnet = as.numeric(res$roc1$auc),
               AUC_rf = as.numeric(res$roc2$auc),
               pvalue = res$p.value,
               method = res$method)
  })

write_csv(delong_rf_glmnet, savepath_data("delong_test", "delong_rf_glmnet.csv"))

```

## Visualize ROC curves for glmnet & RF

```{r}

# Summarize ROC curves for all cancers using RF and glmnet, indicating the p-values obtained using the De-long test
all_roc <- 
  lapply(levels$cancers_12, function(cancer_type) {
    
    roc <- glmnet_res[[cancer_type]]$roc_res 
    
     df<-data.frame(Sensitivity=unlist(roc[2]), 
                    Specificity=unlist(roc[3])) %>% 
    mutate(Group = cancer_type,
           Method = "glmnet",
           AUC = as.numeric(unlist(roc[9])))
     
  }) %>% 
  bind_rows() %>% 
  bind_rows(lapply(levels$cancers_12, function(cancer_type) {
    
    roc <- rf_res[[cancer_type]]$roc_res 
    
    df<-data.frame(Sensitivity=unlist(roc[2]), 
                    Specificity=unlist(roc[3])) %>% 
    mutate(Group = cancer_type,
           Method = "rf",
           AUC = as.numeric(unlist(roc[9])))
     
  }) %>% 
  bind_rows()) %>% 
  mutate(Performance = paste(Group, " - ", Method, ": ", round(AUC,2),sep=""))


aucs <- 
  all_roc %>% 
  select(GROUP = Group, AUC, Method) %>% 
  left_join(levels$cancers_12_mapping, by = "GROUP") %>% 
  mutate(AUC = round(AUC,2)) %>% 
  distinct() %>% 
  mutate(x = 0.95,
         y = case_when(Method == "glmnet" ~ 0.35,
                       Method == "rf" ~ 0.25,
                       T ~ 0.05))

all_roc %>% 
  rename(GROUP = Group) %>% 
  left_join(levels$cancers_12_mapping, by = "GROUP") %>% 
  mutate(Cancer = factor(Cancer, levels = levels$cancers_12_mapping$Cancer)) %>%
  arrange(Cancer,Method) %>% 
  ggplot(aes(x = 1-Specificity, y=Sensitivity, group = Method)) +
  geom_path(aes(color = Method), size=1) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), 
               colour='grey', linetype = 'dotdash') +
  geom_text(data = aucs, aes(x,y,color=Method, label = paste("AUC: ", round(AUC,3), sep = "")),
            hjust = 1,
            size = 3, show.legend = F) +
  geom_text(data = delong_rf_glmnet %>% 
              rename(GROUP = cancer)%>% 
              left_join(levels$cancers_12_mapping, by = "GROUP") , 
            aes(x = 0.7, y = 0.1, label = paste("p-value: ", round(pvalue,2), sep = "")), 
            color = "black", size = 3, inherit.aes = F) +
  themes$main + 
  scale_color_manual(values = c("glmnet" = "#6F70C9", "rf" = "#8ED2C9")) +
  facet_wrap(~as.factor(Cancer)) +
  theme(axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top")

ggsave(savepath("aucs_rf_glmnet_delong.pdf"),
       height = 7,
       width = 7)

```

## Compare protein importance

```{r}

# Extract protein importance scores from glmnet and random forest models
prot_import <- 
  lapply(c(1:length(levels$cancers_12)), function(i){
    model <- glmnet_res[[i]]
    name <- levels$cancers_12[[i]]
    
    model$prot_res %>% 
      mutate(GROUP = name)
  }) %>% 
  bind_rows() %>% 
  mutate(Model = "glmnet") %>% 
  bind_rows(lapply(c(1:length(levels$cancers_12)), function(i){
    model <- rf_res[[i]]
    name <- levels$cancers_12[[i]]
    
    model$prot_res %>% 
      mutate(GROUP = name)
  }) %>% 
  bind_rows() %>% 
  mutate(Model = "rf"))


# Scatterplot - model importance
prot_import %>% 
  left_join(levels$cancers_12_mapping) %>% 
  mutate(Cancer = factor(Cancer, levels = levels$cancers_12_mapping$Cancer)) %>%
  spread(Model, Overall, -1, -3) %>% 
  mutate(avg = (glmnet + rf) / 2) %>% 
  ggplot(aes(glmnet, rf, color = GROUP)) +
  geom_point(alpha = 0.5, show.legend = F) +
  geom_text_repel(data = . %>% 
               group_by(GROUP) %>% 
               top_n(5, avg), 
             aes(label = Protein),
             size = 3,
             color = "black") +
  scale_color_manual(values = palettes$alt_group) +
  facet_wrap(~Cancer, ncol = 4) + #4
  themes$main +
  ylab("random forest") 

ggsave(savepath("protein_importance_rf_glmnet.pdf"),
                height = 6,
                width = 10)

```

## Compare model probabilities

```{r}

# Extract model prediction probabilities from glmnet and random forest models
sample_prob <- 
  lapply(c(1:length(levels$cancers_12)), function(i){
    model <- glmnet_res[[i]]
    name <- levels$cancers_12[[i]]
    
    model$probabilities_test %>% 
      mutate(GROUP = name)
  }) %>% 
  bind_rows() %>% 
  mutate(Model = "glmnet") %>% 
  bind_rows(lapply(c(1:length(levels$cancers_12)), function(i){
    model <- rf_res[[i]]
    name <- levels$cancers_12[[i]]
    
    model$probabilities_test %>% 
      mutate(GROUP = name)
  }) %>% 
  bind_rows() %>% 
  mutate(Model = "rf"))


# Scatterplot - prediction probabilities
sample_prob %>% 
  ungroup() %>% 
  mutate(GROUP = factor(GROUP, levels = levels$cancers_12)) %>% 
  left_join(levels$cancers_12_mapping) %>% 
  mutate(Cancer = factor(Cancer, levels = levels$cancers_12_mapping$Cancer)) %>%
  select(-Probability_control) %>% 
  spread(Model, Probability_cancer, -1,-2,-4) %>% 
  ggplot(aes(glmnet, rf, color = True_class)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~Cancer, ncol = 4) + 
  themes$main +
  ylab("random forest") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0))+
  theme(legend.position = "top")

ggsave(savepath("probabilities_rf_glmnet.pdf"),
                height = 6,
                width = 10)
```

## Save Supplementary file 2

```{r}

model_results <- 
  prot_import %>% 
  mutate(Overall = round(Overall,1)) %>% 
  spread(Model, Overall) %>% 
  left_join(levels$cancers_12_mapping,
            by = "GROUP") %>% 
  mutate(Cancer = factor(Cancer, levels = levels$cancers_12_mapping$Cancer)) %>% 
  mutate(Protein = gsub("`","",Protein))%>% 
  arrange(Cancer) %>% 
  left_join(data %>% 
              select(Protein = Assay,OlinkID, UniProt,Panel) %>% 
              distinct(),
            by = "Protein") %>% 
  select(Cancer,Protein, OlinkID, UniProt, Panel, Overall_importance_rf = rf, Overall_importance_glmnet = glmnet) 

writexl::write_xlsx(model_results, savepath_data("Supplementary_data", "DataS2_prediction_models.xlsx"))

```


# Differential expression

```{r}

# Load results from differential expression analysis (Differential_expression.rmd)
de_all <- 
  readRDS(savepath_data("DE","cancers_sex_controlled_not_adjusted.rds")) 

# Multiple hypothesis testing correction
de_adj <- 
  de_all %>% 
  group_by(Class) %>% 
  do({
    dat <- .
    
    dat %>% 
      DE_adjust()
  })

# Select upregulated proteins
de_up <- 
  de_adj %>% 
  filter(sig == "significant up")

```


# Protein panel

## Select protein panel

```{r}

# Top 3 upregulated proteins per cancer
prot_panel_top3 <- 
  all_prot %>% 
  bind_rows() %>% 
  inner_join(de_up %>% 
               select(Protein = Assay,GROUP = Class),
             by = c("Protein", "GROUP")) %>% 
  rename(Cancer = GROUP) %>%
  group_by(Cancer) %>% 
  filter(Overall > 0) %>% 
  top_n(3, wt = Overall)

# Upregulated proteins with more than 50% importance
prot_panel_50 <- 
  all_prot %>% 
  bind_rows() %>% 
  inner_join(de_up %>% 
               select(Protein = Assay,GROUP = Class),
             by = c("Protein", "GROUP")) %>% 
  rename(Cancer = GROUP) %>% 
  filter(Overall > 50) 

# Protein panel (upregulated proteins with >50% importance, and at least 3 proteins per cancer)
prot_panel_up <- 
  prot_panel_top3 %>% 
  bind_rows(prot_panel_50) %>% 
  distinct()

prot_panel_up %>% 
  pull(Protein) %>% 
  unique() %>% 
  length()

prot_panel_up %>% 
  group_by(Cancer) %>% 
  summarise(n = n_distinct(Protein)) %>% 
  arrange(-n)

saveRDS(prot_panel_up, savepath_data("Panel", "panel_proteins_up.rds"))

```

## Save Supplementary file 3

```{r}

protein_panel_excel <- 
  prot_panel_up %>% 
  mutate(Protein = gsub("`","",Protein))%>% 
  select(GROUP = Cancer, Protein, `Protein Importance` = Overall) %>% 
  left_join(levels$cancers_12_mapping, by = "GROUP") %>% 
  left_join(de_up %>% 
              select(GROUP = Class, Protein = Assay, OlinkID,  `NPX difference` = difference,	`p.adjusted` = p.adjusted, Significance = sig),
            by = c("GROUP","Protein")) %>% 
  ungroup() %>% 
  select(Cancer, Protein, OlinkID, `Protein Importance`, `NPX difference`, `p.adjusted`, Significance) %>% 
  mutate(Significance = str_to_sentence(Significance)) %>% 
  mutate(Cancer = factor(Cancer, levels = levels$cancers_12_mapping$Cancer),
         `Protein Importance` = round(`Protein Importance`,1),
         `NPX difference` = round(`NPX difference`,1)) %>% 
  arrange(Cancer,-`Protein Importance`)


writexl::write_xlsx(protein_panel_excel, savepath_data("Supplementary_data", "DataS3_protein_panel.xlsx"))

```


## Heatmap summarizing expression profiles of panel proteins accross cancers

```{r}

prots_in_panel <- 
  prot_panel_up %>% 
  pull(Protein) %>% 
  unique()

ann_col <- 
  data.frame(X = levels$cancers_12, Cancer = levels$cancers_12) %>% 
  column_to_rownames("X")

ann_row <- 
  prot_panel_up %>% 
  group_by(Protein) %>% 
  top_n(1,Overall) %>% 
  select(-Overall) %>% 
  arrange(Cancer) %>% 
  column_to_rownames("Protein")

sample_order <- 
  data %>% 
  select(Sample) %>% 
  distinct() %>% 
  left_join(metadata %>% 
              select(GROUP,Sample),
            by = "Sample") %>% 
  filter(GROUP %in% levels$cancers_12) %>% 
  arrange(GROUP) 
  
gaps <- 
  ann_row %>% 
  rownames_to_column("Protein") %>% 
  group_by(Cancer) %>% 
  summarise(n = n_distinct(Protein)) %>% 
  mutate(gap = cumsum(n)) %>% 
  head(11) %>% 
  pull(gap)

data %>% 
  left_join(metadata %>% 
              select(GROUP,Sample),
            by = "Sample") %>% 
  filter(GROUP %in% levels$cancers_12,
         Assay %in% prots_in_panel) %>% 
  select(Sample,Assay,NPX) %>% 
  spread(Assay,NPX) %>% 
  column_to_rownames("Sample") %>% 
  scale() %>%
  as.data.frame() %>%
  rownames_to_column("Sample") %>% 
  gather(Assay,NPX,-Sample) %>% 
  left_join(metadata %>% 
              select(GROUP,Sample),
            by = "Sample") %>% 
  group_by(GROUP,Assay) %>% 
  drop_na() %>% 
  summarise(avg_NPX = mean(NPX)) %>% 
  ungroup() %>% 
  mutate(Assay = factor(Assay, levels = rownames(ann_row)),
         GROUP = factor(GROUP, 
                        c("CLL","CVX","PRC",
                          "BRC","ENDC","OVC","CRC",
                          "LUNGC","MYEL","GLIOM",
                          "AML","LYMPH"))) %>% 
  arrange(GROUP,Assay) %>% 
  group_by(Assay) %>% 
  mutate(avg_NPX = scales::rescale(avg_NPX, to = c(0,1))) %>% 
  spread(GROUP,avg_NPX) %>% 
  column_to_rownames("Assay") %>% 
  t() %>% 
  pheatmap(clustering_method = "ward.D2",
           color = palettes$heatmap(100),
           cluster_cols = F,
           annotation_col = ann_row,
           cellwidth = 6,
           cellheight = 6,
           gaps_col = gaps,
           fontsize = 7,
           annotation_colors = list("Cancer" = palettes$alt_group[levels$cancers_12])) %>% 
  as.ggplot()

ggsave(savepath("heatmap_scaled.pdf"),
       height = 5,
       width = 10)

```

## Network summarizing the selected proteins for each cancer

```{r}

set.seed(213)
tbl_graph <- 
  prot_panel_up %>%
  select(Protein, Cancer, Overall) %>% 
  as_tbl_graph() 

tbl_graph %>% 
  activate(nodes) %>% 
  mutate(text_color = case_when(name %in% levels$cancers_12 ~ "white",
                                T ~ "black")) %>%
  create_layout(layout = "nicely") %>%
  ggraph() + 
  geom_edge_link(color = "grey80", width = 1) +
  geom_node_point(data = . %>% 
                    filter(name %in% levels$cancers_12) ,
                  aes(color = name),
                  size = 10,
                  fill ="grey30",
                  shape = 21,
                  stroke = 3) +
  geom_node_point(data = . %>% 
                    filter(!name %in% levels$cancers_12) %>% 
                    left_join(prot_panel_up %>% 
                                select(name = Protein, Overall),
                              by = "name"), 
                  aes(fill = Overall),
                  shape = 21,
                  size = 5,
                  color = "grey80",
                  stroke = 0) +
  scale_color_manual(values = palettes$alt_group[levels$cancers_12], 
                     guide = "none") +
  scale_fill_gradient(high = "grey30",low = "grey80") +
  ggnewscale::new_scale_color() +
  geom_node_text(aes(label = str_wrap(name, width = 10),
                     color = text_color),
                 lineheight = 0.8,
                 size = 2,
                 vjust = 0) +
  scale_color_identity() +
  theme_void() + 
  coord_fixed() 

ggsave(savepath("network_protein_panel.pdf"))

```

## UMAP of all samples based on the panel proteins

```{r}

# Prepare data and use removeBatchEffect from limma to correct for sex effect
meta_data <-
  metadata %>% 
  filter(GROUP %in% levels$cancers_12,
         Sample %in% unique(data$Sample)) %>% 
  select(Sample,GROUP,Sex)

in_dat <- 
  data %>%
  left_join(metadata, by = "Sample") %>% 
  filter(GROUP %in% levels$cancers_12) %>% 
  filter(Assay %in% unique(prot_panel_up$Protein)) %>% 
  select(Sample,Assay,NPX) %>% 
  impute_values(ID = "Sample")

in_dat_matrix <-
  in_dat %>% 
  column_to_rownames("Sample")

in_data_limma <- 
  in_dat_matrix[meta_data$Sample,] %>% 
  t() %>% 
  removeBatchEffect(batch = meta_data$Sex,
                    design = model.matrix(~ GROUP, 
                                          data = meta_data),
                    method = 'ls') 

# Perform PCA + UMAP
data_umap <- 
  in_data_limma %>% 
  t() %>% 
  as.data.frame() %>% 
  do_umap(seed = 13, n_neighbors = 25, n_components = 2, pca = 30, spread = 1)%>% 
  rownames_to_column("Sample") %>% 
  left_join(metadata) %>% 
  left_join(levels$cancers_12_mapping) 

data_umap %>% 
  ggplot(aes(UMAP1, UMAP2, color = GROUP)) +
  geom_point(size = 1, alpha = 0.6) + 
  scale_color_manual(values = palettes$alt_group) +
  scale_fill_manual(values = palettes$alt_group) +
  theme_void() +
  coord_fixed()

ggsave(savepath("umap_25n_30pc_unscaled.pdf"),
       height = 7,
       width = 7)

# Faceted UMAP (highlight one cancer at a time)
tibble(Facet = levels$cancers_12) %>% 
  group_by_all() %>% 
  do({
    data_umap
  }) %>% 
  mutate(color = case_when(Facet == GROUP ~ GROUP,
                           T ~ "Other")) %>% 
  ggplot(aes(UMAP1, UMAP2, color = color)) +
  geom_point(size = 0.2, alpha = 0.6, show.legend = F) +
  scale_color_manual(values = c(palettes$alt_group, "Other" = "grey")) +
  scale_fill_manual(values = c(palettes$alt_group, "Other" = "grey")) +
  facet_wrap(~Facet, nrow = 3)+
  themes$simple+
  coord_fixed()

ggsave(savepath("UMAP_facet.pdf"),
       height = 5,
       width = 10)


```


## Protein boxplots 

```{r}

# Generate boxplot of the top panel protein for each cancer
prots <- 
  prot_panel_up %>% 
  group_by(Cancer) %>% 
  top_n(1,Overall) %>% 
  select(Protein,Cancer)

data %>% 
  select(Sample, Assay, NPX) %>% 
  left_join(metadata %>% 
              select(Sample,GROUP),
            by = "Sample") %>% 
  left_join(levels$cancers_12_mapping %>% 
              rename(Cancer_fac = Cancer),
            by = "GROUP") %>% 
  filter(GROUP %in% levels$cancers_12) %>% 
  filter(Assay %in% prots$Protein) %>%
  left_join(prots, by = c("Assay" = "Protein")) %>% 
  mutate(Assay = factor(Assay, levels = prots$Protein),
         GROUP = factor(GROUP, levels = levels$cancers_12_mapping$GROUP),
         Cancer_fac = factor(Cancer_fac, levels = levels$cancers_12_mapping$Cancer),
         color = case_when(GROUP == Cancer ~ Cancer,
                           T ~ "Other")) %>% 
  ggplot(aes(Cancer_fac, NPX, color = color, fill = color)) +
  geom_quasirandom(size = 0.5, alpha = 0.5, show.legend=F) +
  geom_boxplot(alpha = 0.2, color = "black", outlier.color = NA, show.legend = F) +
  facet_wrap(~Assay, scales = "free_y", nrow = 4) +
  scale_color_manual(values = c(palettes$alt_group, "Other" = "grey")) +
  scale_fill_manual(values = c(palettes$alt_group, "Other" = "grey")) +
  themes$simple +
  theme(axis.text = element_text(size = 8)) 

ggsave(savepath("top_panel_protein_per_cancer.pdf"),
       height = 8,
       width = 8)

# Generate boxplot for all panel proteins
prot_panel_up %>% 
  group_by(Cancer) %>% 
  summarise(n = n_distinct(Protein)) %>% 
  arrange(-n)

prot_panel_up_ex <- 
  prot_panel_up %>% 
  mutate(Cancer = factor(Cancer, levels = levels$cancers_12)) %>% 
  mutate(comb = paste(Protein, " (", Cancer, ")", sep = "")) %>% 
  arrange(Cancer,-Overall)

panel_boxplots <- 
  lapply(prot_panel_up_ex$comb, function(protein) {
    
    current <- 
      prot_panel_up_ex %>% 
      filter(comb == protein)
    
    data %>% 
      select(Sample, Assay, NPX) %>% 
      left_join(metadata %>% 
                  select(Sample,GROUP), 
                by = "Sample") %>% 
      filter(GROUP %in% levels$cancers_12) %>% 
      filter(Assay %in% current$Protein) %>%
      mutate(GROUP = factor(GROUP, levels = levels$cancers_12),
             color = case_when(GROUP == current$Cancer ~ as.character(current$Cancer),
                               T ~ "Other")) %>% 
      ggplot(aes(GROUP, NPX, color = color, fill = color)) +
      geom_quasirandom(size = 0.5, alpha = 0.5, show.legend=F) +
      geom_boxplot(alpha = 0.2, color = "black", outlier.color = NA, show.legend = F) +
      scale_color_manual(values = c(palettes$alt_group, "Other" = "grey")) +
      scale_fill_manual(values = c(palettes$alt_group, "Other" = "grey")) +
      themes$simple +
      theme(axis.text = element_text(size = 8)) +
      ggtitle(current$comb)
    
    
  })

pdf(savepath("all_panel_boxplots.pdf"))
panel_boxplots
dev.off()

```



# Multiclassification 

## All proteins

```{r}

# Multiclassification including all measured proteins (n = 1,463)
multipred_12_all <-
  multiclassification(data = data, 
                      split_data = cancers_split,
                      metadata = metadata,
                      method = "glmnet", 
                      groups = levels$cancers_12, 
                      palette = palettes$alt_group[levels$cancers_12],
                      tune_param = 10, 
                      CV = 5)

saveRDS(multipred_12_all, savepath_data("Models", "multipred_all_proteins.rds"))

```


## Panel

```{r}

# Multiclassification including all proteins included in the panel
multipred_12_panel <-
  multiclassification(data = data, 
                      split_data = cancers_split,
                      metadata = metadata,
                      select_proteins = unique(prot_panel_up$Protein),
                      method = "glmnet", 
                      groups = levels$cancers_12, 
                      palette = palettes$alt_group[levels$cancers_12],
                      tune_param = 10, 
                      CV = 5)

saveRDS(multipred_12_panel, savepath_data("Models", "multipred_panel_proteins.rds"))

```


## Top 3

```{r}

# Multiclassification including the top3 proteins in the panel for each cancer (n = 36)
top_3 <- 
  prot_panel_up %>% 
  group_by(Cancer) %>% 
  top_n(3, Overall)

multipred_top3_panel <-
  multiclassification(data = data, 
                      split_data = cancers_split,
                      metadata = metadata,
                      select_proteins = unique(top_3$Protein),
                      method = "glmnet", 
                      groups = levels$cancers_12, 
                      palette = palettes$alt_group[levels$cancers_12],
                      tune_param = 10, 
                      CV = 5)

saveRDS(multipred_top3_panel, savepath_data("models", "multipred_top3_cancer_proteins.rds"))

```

## Top 1

```{r}

# Multiclassification including the top3 proteins in the panel for each cancer (n = 12)
top_1 <- 
  prot_panel_up %>% 
  group_by(Cancer) %>% 
  top_n(1, Overall)

multipred_top1_panel <-
  multiclassification(data = data, 
                      split_data = cancers_split,
                      metadata = metadata,
                      select_proteins = unique(top_1$Protein),
                      method = "glmnet", 
                      groups = levels$cancers_12, 
                      palette = palettes$alt_group[levels$cancers_12],
                      tune_param = 10, 
                      CV = 5)

saveRDS(multipred_top1_panel, savepath_data("Models", "multipred_top1_cancer_proteins.rds"))

```


## ROC comparison

```{r}

# Extract the ROC results from all multiclassifications
all_roc <- 
  multipred_top1_panel$roc_df %>% 
  select(-Method) %>% 
  mutate(Method = "top 1",
         Performance = paste(Group, " - ", Method, ": ", round(AUC,2),sep="")) %>% 
  bind_rows( multipred_top3_panel$roc_df %>% 
               select(-Method) %>% 
               mutate(Method = "top 3",
                      Performance = paste(Group, " - ", Method, ": ", round(AUC,2),sep=""))) %>% 
bind_rows(multipred_12_panel$roc_df %>% 
            select(-Method) %>% 
            mutate(Method = "protein panel",
                   Performance = paste(Group, " - ", Method, ": ", round(AUC,2),sep=""))) %>% 
  bind_rows(multipred_12_all$roc_df %>% 
              select(-Method) %>% 
              mutate(Method = "all proteins",
                     Performance = paste(Group, " - ", Method, ": ", round(AUC,2),sep="")))

aucs <- 
  all_roc %>% 
  rename(GROUP = Group) %>% 
  left_join(levels$cancers_12_mapping, by = "GROUP") %>% 
  select(Cancer, AUC, Method) %>% 
  mutate(AUC = round(AUC,2)) %>% 
  distinct() %>% 
  mutate(x = 0.75,
         y = case_when(Method == "all proteins" ~ 0.35,
                       Method == "protein panel" ~ 0.25,
                       Method == "top 3" ~ 0.15,
                       T ~ 0.05))

# Visualize all ROC 
all_roc %>%
  rename(GROUP = Group) %>% 
  left_join(levels$cancers_12_mapping, by = "GROUP") %>% 
  mutate(Cancer = factor(Cancer, levels = levels$cancers_12_mapping$Cancer),
         Method = factor(Method,levels = c("all proteins", "protein panel", "top 3", "top 1"))) %>%
  arrange(Cancer,Method) %>% 
  ggplot(aes(x = 1-Specificity, y=Sensitivity, group = Method)) +
  geom_path(aes(color = Method), size=1) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), 
               colour='grey', linetype = 'dotdash') +
  geom_text(data = aucs, aes(x,y,color=Method, label = paste("AUC: ", round(AUC,3),sep = "")), size = 3, show.legend = F) +
  themes$main + 
  scale_color_manual(values = palettes$auc) +
  facet_wrap(~as.factor(Cancer)) +
  theme(axis.text = element_text(size = 8),
        legend.position = "top")

ggsave(savepath("multipred_aucs_top.pdf"),
       height = 7,
       width = 8)

```


# Classification Cancer-Wellness

## Data split

```{r}

# Prepare Wellness data
wellness_data_complete <- 
  wellness_data %>%  
  left_join(wellness_meta, by = "Sample") %>% 
  mutate(Study = "Wellness")

wellness_dat <- 
  wellness_data_complete %>% 
  select(Sample, GROUP) %>% 
  distinct()

# Split Wellness data into train and test set
set.seed(213)
train <- createDataPartition(wellness_dat$GROUP,p=0.7,list=FALSE)
train_data <- 
  wellness_dat[train,] %>% 
  mutate(set = "train")
test_data <- wellness_dat[-train,] %>% 
  mutate(set = "test")

wellness_split <- 
  train_data %>% 
  bind_rows(test_data)

final_split <- 
  cancers_split %>% 
  bind_rows(wellness_split)

# Combined Cancer & Wellness metadata
all_meta_w <- 
  metadata %>% 
  filter(GROUP %in% levels$cancers_12,
         Sample %in% data$Sample) %>% 
  select(Sample,GROUP) %>% 
  mutate(Study = "UCAN") %>% 
  bind_rows(wellness_dat %>% 
              mutate(Study = "Wellness"))

```


## Limma correction

```{r}

# Batch correction using limma
ucan_wellness <- 
  data %>% 
  mutate(Study = "UCAN") %>% 
  left_join(metadata %>% 
              select(Sample, GROUP, Sex),
            by = "Sample") %>% 
  filter(GROUP %in% levels$cancers_12) %>% 
  bind_rows(wellness_data_complete)

to_limma_u_w <- 
  ucan_wellness %>% 
  left_join(all_meta_w, by = c("Sample","GROUP","Study")) %>% 
  filter(Study %in% c("UCAN","Wellness")) %>% 
  select(-Study) %>% 
  mutate(Sample = factor(Sample, levels = all_meta_w$Sample)) %>% 
  arrange(Sample) %>% 
  select(Sample,Assay, NPX) %>% 
  spread(Sample,NPX) %>% 
  column_to_rownames("Assay") %>% 
  as.matrix()

batch <- 
  all_meta_w %>% 
  filter(Sample %in% colnames(to_limma_u_w)) %>%
  mutate(batch =  Study) %>% 
  pull(batch)

corrected_limma_u_w <- 
  removeBatchEffect(to_limma_u_w, batch)

all_data_limma_u_w <- 
  corrected_limma_u_w %>%
  t() %>% 
  as_tibble(rownames = "Sample") %>% 
  gather(Assay,NPX, -Sample) %>% 
  left_join(all_meta_w) %>% 
  filter(!is.na(NPX))%>% 
  mutate(correction = "limma")


saveRDS(all_meta_w, savepath_data("Wellness_comparison", "all_meta_w.rds"))
saveRDS(final_split, savepath_data("Wellness_comparison", "final_split.rds"))
saveRDS(wellness_data_complete, savepath_data("Wellness_comparison", "wellness_data_complete.rds"))
saveRDS(all_data_limma_u_w, savepath_data("Wellness_comparison", "all_data_limma_u_w.rds"))

```


## Wellness - Cancer classification (glmnet)

```{r}

wellness_limma <- 
  lapply(levels$cancers_12, function(cancer) {
    
    
    if(cancer %in% c("BRC","ENDC","CVX","OVC")) {
      all_data_split <- 
        cancers_split %>% 
        filter(GROUP  == cancer) %>% 
        bind_rows(wellness_split %>% 
                    left_join(wellness_meta %>% 
                                select(Sample,Sex),
                              by = "Sample") %>% 
                    filter(Sex == "f") %>% 
                    select(-Sex))
      
    } else if (cancer == "PRC") {
      all_data_split <- 
        cancers_split %>% 
        filter(GROUP  == cancer) %>% 
        bind_rows(wellness_split %>% 
                    left_join(wellness_meta %>% 
                                select(Sample,Sex),
                              by = "Sample") %>% 
                    group_by(set,Sex) %>% 
                    filter(Sex == "m") %>% 
                    select(-Sex))
    } else {
      all_data_split <- 
        cancers_split %>% 
        filter(GROUP  == cancer) %>% 
        bind_rows(wellness_split)
      
    }
    
    cancer_proteins <- 
      prot_panel_up %>% 
      filter(Cancer == cancer) %>% 
      pull(Protein)
    
    limma_dat <- 
      all_data_limma_u_w %>% select(Sample,Assay,NPX)
    
    if(length(cancer_proteins) > 1) {
      disease_classifier(cancer_type = cancer,
                         method = "glmnet", 
                         data = limma_dat,
                         split_data = all_data_split,
                         metadata = all_meta_w, 
                         select_proteins = cancer_proteins, 
                         tune_param = 10, 
                         CV = 5)
    } else {
      tibble()
    }

  }) %>% 
  set_names(levels$cancers_12)

saveRDS(wellness_limma, savepath_data("Models", "wellness_limma_models.rds"))

```


### Summarize classification results

```{r}

# Generate a plot with the prediction probabilities and ROC results for all the Cancer-Wellness prediction models
wellness_cancer_plots <- 
  lapply(levels$cancers_12, function(cancer) {
    
    cancer_name <- 
      levels$cancers_12_mapping %>% 
      filter(GROUP == cancer) %>% 
      pull(Cancer)
    
    if(is_tibble(wellness_limma[[cancer]])) {
      
      ggplot() + theme_void() # The cancer had 1 or 0 proteins in the panel
      
    } else {
      all_prob <- 
        data.frame(response = wellness_limma[[cancer]]$roc_res$response, 
                   predicted = wellness_limma[[cancer]]$roc_res$predictor) %>% 
        mutate(True_class = case_when(response == "0_Control" ~ "Wellness",
                                      T ~ cancer_name)) %>% 
        rename(Cancer_probability = predicted) %>% 
        ggplot(aes(True_class,Cancer_probability, fill = True_class, color = True_class)) +
        geom_quasirandom(alpha = 0.7,show.legend = F) +
        geom_violin( color = "grey20", alpha = 0.3, show.legend = F, scale = "width")+ 
        stat_summary(fun = "mean",
                     geom = "crossbar", 
                     width = 0.5,
                     color = "grey20",
                     show.legend = F)+
        scale_color_manual(values = c(palettes$alt_group[[cancer]],"grey"))+
        scale_fill_manual(values = c(palettes$alt_group[[cancer]],"grey"))+
        themes$main 
      
      df<-
        data.frame(y=unlist(wellness_limma[[cancer]]$roc_res[2]), x=unlist(wellness_limma[[cancer]]$roc_res[3])) %>% 
        mutate(GROUP = cancer)
      
      all_roc <- 
        df %>% 
        mutate(x = 1-x) %>% 
        arrange(y) %>% 
        ggplot(aes(x, y)) + 
        geom_line(size = 1, color = palettes$alt_group[[cancer]]) + 
        geom_abline(intercept=0, slope=1, linetype="dashed") +
        xlab("1-Specificity") +  #False Positive rate
        ylab("Sensitivity") + #True Positive rate
        themes$main +
        annotate("text", x = .75, y = .25, label = paste("AUC",wellness_limma[[cancer]]$roc_res[[9]] %>% round(3)))
      
      (all_prob/ all_roc ) +
        plot_annotation(title = paste(cancer, " - all samples", sep =""))
    }
    
  })

(wellness_cancer_plots[[1]] | wellness_cancer_plots[[2]] | wellness_cancer_plots[[3]] | wellness_cancer_plots[[4]])  /
  (wellness_cancer_plots[[5]] | wellness_cancer_plots[[6]] | wellness_cancer_plots[[7]] | wellness_cancer_plots[[8]]) /
  (wellness_cancer_plots[[9]] | wellness_cancer_plots[[10]] | wellness_cancer_plots[[11]] | wellness_cancer_plots[[12]])

ggsave(savepath("wellness_classification_results.pdf"),
       height = 13,
       width = 10)

```


## Wellness - Cancer Stage 1 classification (glmnet)

```{r}

# Apply the prediction model Cancer-Wellness to stage 1 patients only (for colorectal and lung cancer)

crc_lung_stage1 <- 
  lapply(c("CRC","LUNGC"), function(cancer) {
    
   # cancer <- "CRC"
    all_data_split <- 
        cancers_split %>% 
        filter(GROUP  == cancer) %>% 
        bind_rows(wellness_split)
        
    cancer_proteins <- 
      prot_panel_up %>% 
      filter(Cancer == cancer) %>% 
      pull(Protein)
    
    disease_classifier_stage1(cancer_type = cancer,
                       method = "glmnet", 
                       data = all_data_limma_u_w %>% select(Sample,Assay,NPX), #all_data,
                       split_data = all_data_split,
                       metadata = all_meta_w,
                       metadata_stage = metadata, 
                       select_proteins = cancer_proteins, 
                       tune_param = 10, 
                       CV = 5)
  }) %>% 
  set_names(c("CRC","LUNGC"))

saveRDS(crc_lung_stage1, savepath_data("Models", "glmnet_res_wellness_crc_lung_stage1.rds"))

```

### Summaryize classification results

```{r}

# Summarize results for colorectal cancer
crc_all_prob <- 
  data.frame(response = crc_lung_stage1$CRC$roc_res$response, predicted = crc_lung_stage1$CRC$roc_res$predictor) %>% 
  mutate(True_class = case_when(response == "0_Control" ~ "Wellness",
                                T ~ "CRC")) %>% 
  rename(Cancer_probability = predicted) %>% 
  ggplot(aes(True_class,Cancer_probability, fill = True_class, color = True_class)) +
  geom_quasirandom(alpha = 0.7,show.legend = F) +
  geom_violin( color = "grey20", alpha = 0.3, show.legend = F, scale = "width")+ 
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               color = "grey20",
               show.legend = F)+
  scale_color_manual(values = c(palettes$alt_group["CRC"],"Wellness"="grey"))+
  scale_fill_manual(values = c(palettes$alt_group["CRC"],"Wellness"="grey"))+
  themes$main 

df<-
  data.frame(y=unlist(crc_lung_stage1$CRC$roc_res[2]), x=unlist(crc_lung_stage1$CRC$roc_res[3])) %>% 
  mutate(GROUP = "CRC")

crc_all_roc <- 
  df %>% 
  mutate(x = 1-x) %>% 
  arrange(y) %>% 
  ggplot(aes(x, y)) + 
  geom_line(size = 1, color = palettes$alt_group[["CRC"]]) + 
  geom_abline(intercept=0, slope=1, linetype="dashed") +
  xlab("False Positive rate") + 
  ylab("True Positive rate") +
  themes$main +
  annotate("text", x = .75, y = .25, label = paste("AUC",crc_lung_stage1$CRC$roc_res[[9]] %>% round(3)))

crc_all_samples <- 
  (crc_all_roc | crc_all_prob) +
  plot_annotation(title = "CRC - all samples")

crc_s1_prob <- 
  data.frame(response = crc_lung_stage1$CRC$roc_res_s1$response, predicted = crc_lung_stage1$CRC$roc_res_s1$predictor) %>% 
  mutate(True_class = case_when(response == "0_Control" ~ "Wellness",
                                T ~ "CRC Stage 1")) %>% 
  rename(Cancer_probability = predicted) %>% 
  ggplot(aes(True_class,Cancer_probability, fill = True_class, color = True_class)) +
  geom_quasirandom(alpha = 0.7,show.legend = F) +
  geom_violin( color = "grey20", alpha = 0.3, show.legend = F, scale = "width")+ 
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               color = "grey20",
               show.legend = F)+
  scale_color_manual(values = c("CRC Stage 1" = palettes$stage[["1"]],"Wellness"="grey")) +
  scale_fill_manual(values = c("CRC Stage 1" = palettes$stage[["1"]],"Wellness"="grey")) +
  themes$main 

df_s1_crc <-
  data.frame(y=unlist(crc_lung_stage1$CRC$roc_res_s1[2]), x=unlist(crc_lung_stage1$CRC$roc_res_s1[3])) %>% 
  mutate(GROUP = "CRC")

crc_s1_roc <- 
  df_s1_crc %>% 
  mutate(x = 1-x) %>% 
  arrange(y) %>% 
  ggplot(aes(x, y)) + 
  geom_line(size = 1, color = palettes$stage[["1"]]) + 
  geom_abline(intercept=0, slope=1, linetype="dashed") +
  xlab("1-Specificity") + 
  ylab("Sensitivity") +
  themes$main +
  annotate("text", x = .75, y = .25, label = paste("AUC",crc_lung_stage1$CRC$roc_res_s1[[9]] %>% round(3)))

crc_s1_samples <- 
  (crc_s1_roc | crc_s1_prob) +
  plot_annotation(title = "CRC - stage 1 samples")

crc_all_samples /
  crc_s1_samples

ggsave(savepath("CRC_all_stages.pdf"))

# Sumamrize results for lung cancer
lungc_all_prob <- 
  data.frame(response = crc_lung_stage1$LUNGC$roc_res$response, predicted = crc_lung_stage1$LUNGC$roc_res$predictor) %>% 
  mutate(True_class = case_when(response == "0_Control" ~ "Wellness",
                                T ~ "LUNGC")) %>% 
  rename(Cancer_probability = predicted) %>% 
  ggplot(aes(True_class,Cancer_probability, fill = True_class, color = True_class)) +
  geom_quasirandom(alpha = 0.7,show.legend = F) +
  geom_violin( color = "grey20", alpha = 0.3, show.legend = F, scale = "width")+ 
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               color = "grey20",
               show.legend = F)+
  scale_color_manual(values = c(palettes$alt_group["LUNGC"],"Wellness"="grey"))+
  scale_fill_manual(values = c(palettes$alt_group["LUNGC"],"Wellness"="grey"))+
  themes$main 

df_lung<-
  data.frame(y=unlist(crc_lung_stage1$LUNGC$roc_res[2]), x=unlist(crc_lung_stage1$LUNGC$roc_res[3])) %>% 
  mutate(GROUP = "LUNGC")

lungc_all_roc <- 
  df_lung %>% 
  mutate(x = 1-x) %>% 
  arrange(y) %>% 
  ggplot(aes(x, y)) + 
  geom_line(size = 1, color = palettes$alt_group[["LUNGC"]]) + 
  geom_abline(intercept=0, slope=1, linetype="dashed") +
  xlab("1-Specificity") +  
  ylab("Sensitivity") +
  themes$main +
  annotate("text", x = .75, y = .25, label = paste("AUC",crc_lung_stage1$LUNGC$roc_res[[9]] %>% round(3)))

lungc_all_samples <- 
  (lungc_all_roc | lungc_all_prob) +
  plot_annotation(title = "LUNGC - all samples")

lungc_s1_prob <- 
  data.frame(response = crc_lung_stage1$LUNGC$roc_res_s1$response, predicted = crc_lung_stage1$LUNGC$roc_res_s1$predictor) %>% 
  mutate(True_class = case_when(response == "0_Control" ~ "Wellness",
                                T ~ "LUNGC Stage 1")) %>% 
  rename(Cancer_probability = predicted) %>% 
  ggplot(aes(True_class,Cancer_probability, fill = True_class, color = True_class)) +
  geom_quasirandom(alpha = 0.7,show.legend = F) +
  geom_violin( color = "grey20", alpha = 0.3, show.legend = F, scale = "width")+ 
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               color = "grey20",
               show.legend = F)+
  scale_color_manual(values = c("LUNGC Stage 1" = palettes$stage[["1"]],"Wellness"="grey")) +
  scale_fill_manual(values = c("LUNGC Stage 1" = palettes$stage[["1"]],"Wellness"="grey")) +
  themes$main 

df_s1_lungc <-
  data.frame(y=unlist(crc_lung_stage1$LUNGC$roc_res_s1[2]), x=unlist(crc_lung_stage1$LUNGC$roc_res_s1[3])) %>% 
  mutate(GROUP = "LUNGC")

lungc_s1_roc <- 
  df_s1_lungc %>% 
  mutate(x = 1-x) %>% 
  arrange(y) %>% 
  ggplot(aes(x, y)) + 
  geom_line(size = 1, color = palettes$stage[["1"]]) + 
  geom_abline(intercept=0, slope=1, linetype="dashed") +
  xlab("1-Specificity") +  #False Positive rate
  ylab("Sensitivity") + #True Positive rate
  #theme_bw() +
  themes$main +
  #theme(panel.grid.major = element_blank(), 
  #       panel.grid.minor = element_blank(),
  #      axis.title = element_text(size = 8)) +
  annotate("text", x = .75, y = .25, label = paste("AUC",crc_lung_stage1$LUNGC$roc_res_s1[[9]] %>% round(3)))

lungc_s1_samples <- 
  (lungc_s1_roc | lungc_s1_prob) +
  plot_annotation(title = "LUNGC - stage 1 samples")

lungc_all_samples /
  lungc_s1_samples

ggsave(savepath("LUNGC_all_stages.pdf"))

```

### De-long test

```{r}

delong_stage1 <- 
  data.frame(cancer = c("CRC","LUNGC")) %>% 
  group_by(cancer) %>% 
  do({
    
    c <- .$cancer
    
    res <- 
      pROC::roc.test(crc_lung_stage1[[c]]$roc_res,
                     crc_lung_stage1[[c]]$roc_res_s1, 
                     method = "delong", 
                     paired = FALSE,
                     boot.n=10000)
    
    data.frame(AUC_all = as.numeric(res$roc1$auc),
               AUC_stage1 = as.numeric(res$roc2$auc),
               pvalue = res$p.value,
               #CI = res$conf.int,
               method = res$method)
  })

write_csv(delong_stage1, savepath_data("delong_test", "delong_all_stage1.csv"))

# Visualize
all_roc <- 
  lapply(c("CRC","LUNGC"), function(cancer_type) {
    
    roc <- crc_lung_stage1[[cancer_type]]$roc_res 
    
     df<-data.frame(Sensitivity=unlist(roc[2]), 
                    Specificity=unlist(roc[3])) %>% 
    mutate(Group = cancer_type,
           Method = "all",
           AUC = as.numeric(unlist(roc[9])))
     
  }) %>% 
  bind_rows() %>% 
  bind_rows(lapply(c("CRC","LUNGC"), function(cancer_type) {
    
    roc <- crc_lung_stage1[[cancer_type]]$roc_res_s1 
    
    df<-data.frame(Sensitivity=unlist(roc[2]), 
                    Specificity=unlist(roc[3])) %>% 
    mutate(Group = cancer_type,
           Method = "stage1",
           AUC = as.numeric(unlist(roc[9])))
     
  }) %>% 
  bind_rows()) %>% 
  mutate(Performance = paste(Group, " - ", Method, ": ", round(AUC,2),sep=""))

aucs <- 
  all_roc %>% 
  select(Group, AUC, Method) %>% 
  mutate(AUC = round(AUC,2)) %>% 
  distinct() %>% 
  mutate(x = 0.95,
         y = case_when(Method == "all" ~ 0.35,
                       Method == "stage1" ~ 0.25,
                       T ~ 0.05))

all_roc %>% 
  arrange(Group,Method) %>% 
  ggplot(aes(x = 1-Specificity, y=Sensitivity, group = Method)) +
  geom_path(aes(color = Method), size=1) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), 
               colour='grey', linetype = 'dotdash') +
  geom_text(data = aucs, aes(x,y,color=Method, label = paste("AUC: ", round(AUC,3), sep ="")),
            hjust = 1,
            size = 3, show.legend = F) +
  geom_text(data = delong_stage1 %>% rename(Group = cancer), 
            aes(x = 0.95, y = 0.1, label = paste("p-value: ", round(pvalue,2), sep = "")),
            hjust = 1,
            color = "black", size = 3, inherit.aes = F) +
  themes$main + 
  scale_color_manual(values = c("all" = "grey", "stage1" = palettes$stage[["1"]])) +
  facet_wrap(~as.factor(Group)) +
  theme(axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45))

ggsave(savepath("aucs_all_stage1_delong.pdf"),
       height = 3,
       width = 6)

```



# Session info 

```{r}

writeLines(capture.output(sessionInfo()), savepath_data("Sessions", "sessionInfo_prediction.txt"))

```

